\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-26 paper.tex
Move Wiuf & Hein caveat
I think it distracts the reader to put the caveat about "not a standard simulated ARG" at the end of the section. This is my attempt to shorten the text and put something up front instead. I think it makes the section end on a stronger note.

Any eARG can be converted to a sample-resolved gARG using the same machinery,
and the two-step process is illustrated in Fig.~\ref{fig-ancestry-resolution}.
The first step is to take the input eARG (Fig.~\ref{fig-ancestry-resolution}A),
duplicate its graph topology, and then add inheritance annotations
to each of the gARG's edges (Fig.~\ref{fig-ancestry-resolution}B) as follows.
If a given node is a common ancestor event, we annotate the single
outbound edge with the interval $[0,L)$, for a genome of length $L$. If the
node is a recombination event with a breakpoint $x$, we annotate the two
outbound edges respectively with the intervals $[0, x)$ and $[x, L)$. These
inheritance interval annotations are clearly in one-to-one correspondence with
the information in the input eARG. They are also analogous to the
inheritance intervals we get on the edges in a prospective gARG
produced by a forwards-time simulation, which are concerned with recording
the local genetic relationship between a parent and child genome and not
necessarily minimal in terms of the resolved ancestral material of a sample.
Thus, the final step is to use the ``simplify'' algorithm to perform this
resolution and produce the sample-resolved
gARG (Fig.~\ref{fig-ancestry-resolution}C).

@@

Note that the \citet{wiuf1999recombination} eARG
in Fig.~\ref{fig-ancestry-resolution} is not particularly
representative, because inference or simulation methods will usually
only generate ARGs containing nodes and edges ancestral to the sample
(see the discussion of the ``Big ARG'' stochastic process in
Appendix~\ref{sec-big-and-little-arg}, however).
Nonetheless, it is an instructive example from the literature which highlights several
important properties of ARGs, and the general point about
the need to resolve ancestral material ``on the fly'' for eARG traversals
holds.

>>>>

Any eARG can be converted to a sample-resolved gARG using the same machinery,
and the two-step process is shown in Fig.~\ref{fig-ancestry-resolution}.
In order to show a graph with non-ancestral material, rather than a simulated
Little ARG (Appendix~\ref{sec-big-and-little-arg}), we use a
classic example from the literature
\citet{wiuf1999recombination}, hand-crafted to
highlight several important ARG properties. The first conversion
step is to take the input eARG 
Fig.~\ref{fig-ancestry-resolution}A),
duplicate its graph topology, and then add inheritance annotations
to each of the gARG's edges (Fig.~\ref{fig-ancestry-resolution}B) as follows.
If a given node is a common ancestor event, we annotate the single
outbound edge with the interval $[0,L)$, for a genome of length $L$. If the
node is a recombination event with a breakpoint $x$, we annotate the two
outbound edges respectively with the intervals $[0, x)$ and $[x, L)$. These
inheritance interval annotations are clearly in one-to-one correspondence with
the information in the input eARG. They are also analogous to the
inheritance intervals we get on the edges in a prospective gARG
produced by a forwards-time simulation, which are concerned with recording
the local genetic relationship between a parent and child genome and not
necessarily minimal in terms of the resolved ancestral material of a sample.
Thus, the final step is to use the ``simplify'' algorithm to perform this
resolution and produce the sample-resolved
gARG (Fig.~\ref{fig-ancestry-resolution}C).

@@
